Mixed language processing increases cross-language phonetic transfer in Bengali-English Bilinguals

### 2021-06-08 draft_glossa-psych/glossa-psych_AM-ID_manuscript.tex
edits

\begin{abstract}
This study investigates the phonetic outcome of mixed-language production in a group of highly proficient bilingual speakers of Bengali and English, to better understand the nature of cross-language phonetic interaction. Existing research has largely focused on a small set of phonologically related languages and used voice onset time (VOT) to measure transfer. Results show that transfer is mediated by both language-specific features and linguistic experience of participants, pointing to an urgent need for diversifying the range of languages and populations studied. We compare productions from a yet-unstudied population in two conditions—unilingual L2 (English) and mixed (English and Bengali). The spectral properties of two L2 vowel categories— \nt{\ae}(common to L1 and L2) and \nt{2} (absent in L1) are examined as sites for transfer. To understand how this unfolds dynamically, we measure F1 and F2 at five points over the course of articulating the vowel. We find a dynamic change in both vowels during mixed production, with \nt{2} shifting towards a related L1 category. We interpret this as a temporary increase in cross-language phonetic influence during mixed-language use, leading to a shift towards L1 norms. asymmetrical shifts in the two categories, and no difference in transfer patterns between the two paradigms. 
\end{abstract}

@@

Bi/multilingual speakers can distinguish between the phonetic norms of their languages and maintain separate sound categories for each language \citep{caramazza1973acquisition,macleod2010impact,bosch2003simultaneous}. However, these categories are not autonomous-- they influence each other across languages in both perception and production \citep[e.g.][]{flege1995second,fowler2008cross,flege2002assessing}, and the nature of such interaction provides crucial insights into how language `systems' are cognitively represented and processed. This study reports on the phonetic outcomes of mixed-language production in proficient bilingual speakers of Bengali and English.\\%Given that a majority of the world's population is multilingual, this is a central concern for any realistic theory of language.
%show cross-language-- rephrase

%keep just the types here, and put the next 2 lines in the causes&effects section?
Cross-language influence at the level of sounds can be studied broadly in two kinds of conditions:
(i) While a bilingual speaker is operating in any one of their languages, by comparing bilingual speech to monolingual norms \citep[e.g.][]{guion2003vowel,caramazza1973acquisition,flege1987production}; %These studies often compare the speech of bilinguals to monolingual norms, and view the influence as changes to long-term memory representations as a result of acquiring an L2 \citep{guion2003vowel,caramazza1973acquisition,flege1987production}; 
(ii) When both languages of a bilingual speaker are co-activated, by comparing speakers' mixed-language speech to their own norms while using a single language \citep[e.g.][]{grosjean1994going, bullock2009trying,elias2017effects, simonet2014phonetic}. %These compare productions during mixed-language use to participants' own productions while using a single language. The resulting influence is variously thought to involve short-term memory, online processing costs, language mode, and context-awareness. 
The present study is concerned with the latter. In production, researchers have variously referred to this kind of influence as transfer, drift, accommodation, and interference-- these terms are used interchangeably here to indicate any interaction between two sets of phonetic norms.\\

Existing research on the phonetic effects of mixed-language production has largely focused on a few pairs of phonologically related languages. A majority of these studies use temporal properties of consonants (in particular, voice onset time or VOT) to measure transfer. However, the reported results vary greatly across studies, and appear to be contingent upon both language-specific features and language experience of the participants. This suggests that data from a wider variety of
%elaborate on 'great variation'-- differences between language pairs, plus sociolinguistic situation (effects of language experience etc)-- clearer. 'Types of features'--separate line?- Brief mention - discursive factors in VOT. 
populations, language pairs,  and phonetic features needs to be considered in order to make meaningful generalizations. There is no work yet on short-term phonetic transfer in the Indian subcontinent, or in any Indo-Aryan language. Nevertheless, widespread multilingualism in this part of the world suggests that phonetic
%line--rephrase
behavior in these populations can be particularly valuable towards understanding the nature of such cross-language interactions, as they are likely to reflect real-world experience with mixed-language processing.\\

The present study examines phonetic transfer between Bengali and English in a group of highly proficient bilingual speakers in India. We measure spectral properties (F1 and F2) of two English vowels to ascertain if L1 influence on L2 increases during mixed-language use, relative to a participant's unilingual baseline production of L2. Mixed-language data is elicited in two switching paradigms: code-switching and cued picture-naming. We compare these results to assess if differences between the paradigms independently influence the outcome of phonetic interaction. The results demonstrate a dynamic effect of transfer on L2 vowels during mixed-language production, but no difference in the extent of transfer between the two paradigms. Findings are discussed in light of recent proposals about asymmetries in short-term phonetic interaction and the role of connected speech in introducing discursive factors to transfer studies.\\

The rest of this section is structured as follows: \ref{bengali_english_in_india} provides background information on Bengali and Indian English focusing on vowel systems. \ref{causes} discusses two potential sources of transfer during language processing-- global co-activation of multiple languages vs the local act of switching between languages. Given the highly multilingual setting that characterizes this population, we note that the former is expected to be a constant feature and therefore any observed phonetic interaction is better understood as resulting from the latter. The design of the present study is motivated on the basis of this discussion. A logical consequence of this model is that such phonetic interaction must be highly localized. \ref{duration} elaborates on this by reviewing literature on the duration and `reversiblity' of transfer effects. A recurring pattern that emerges from existing research on short-term phonetic transfer is that the phonetic influence does not affect both languages of the bilingual speaker equally, although the nature of these reported differences varies greatly across studies. In \ref{asymmetries}, we review some of these findings, focusing on proposed sources of asymmetries, in order to highlight the factors that seem to mediate such phonetic interaction. We argue that these results point to an urgent need for expanding the range of language pairs and sound categories examined in order to make meaningful generalizations about the nature of short-term phonetic transfer. Following from this, \ref{asymmetry between sounds} discusses reported asymmetries between different sound categories of a single language, and develops the hypotheses of the present study on the basis of these. In \ref{paradigms}, we motivate a secondary aim of this study: to verify the postulated differences between two paradigms for eliciting short-term phonetic interaction in an experimental setup, namely code-switching and cued picture-naming. \ref{questions_and_hypotheses} summarizes the research questions and hypotheses.

@@

	o-ta \quad Ek-ta \quad kalo [target word]\\
	that-CL one-CL black [target word]\\
	that is a black [target word]\\

>>>>

\begin{abstract}
This study investigates the phonetic outcome of mixed-language production in a group of highly proficient bilingual speakers of Bengali and English, to better understand the nature of cross-language phonetic interaction. Existing research has largely focused on a small set of phonologically related languages and used voice onset time (VOT) to measure phonetic transfer. Results show that this interaction is mediated by both language-specific features and linguistic experience of participants, necessitating diversification in the range of languages and populations studied. We compare productions from a yet-unstudied population in two conditions-- unilingual L2 (English) and mixed (Bengali and English). The spectral properties of two L2 vowel categories— \nt{\ae}(common to L1 and L2) and \nt{2} (absent in L1) are examined as sites for transfer. To understand how this unfolds dynamically, we measure F1 and F2 at five points over the course of articulating the vowel. We find dynamic and  asymmetrical changes in both vowels during mixed production, with \nt{2} shifting towards a related L1 category. We interpret this as a temporary increase in cross-language phonetic influence during mixed-language use, leading to a shift towards L1 norms. Mixed language is elicited in two switching paradigms: cued picture-naming and code-switching, to assess if differences between the paradigms independently influence the outcome of phonetic interaction. Results reveal identical transfer patterns across the two paradigms. Findings are discussed in light of recent proposals about asymmetries in short-term phonetic interaction and the role of connected speech in introducing discursive factors to transfer studies. 
\end{abstract}

@@

Bi/multilingual speakers can distinguish between the phonetic norms of their languages and maintain separate sound categories for each language \citep{caramazza1973acquisition,macleod2010impact,bosch2003simultaneous}. However, these categories are not autonomous-- they influence each other across languages in both perception and production \citep[e.g.][]{flege1995second,fowler2008cross,flege2002assessing}, and the nature of such interaction provides crucial insights into how language `systems' are cognitively represented and processed. Based on the productions of proficient bilingual speakers of Bengali and English, this study reports a temporary increase in cross-language phonetic interaction during mixed-language use.\\%Given that a majority of the world's population is multilingual, this is a central concern for any realistic theory of language.
%show cross-language-- rephrase

%keep just the types here, and put the next 2 lines in the causes&effects section?
Cross-language influence at the level of sounds can be studied broadly in two kinds of conditions:
(i) While a bilingual speaker is operating in any one of their languages, by comparing bilingual speech to monolingual norms \citep[e.g.][]{guion2003vowel,caramazza1973acquisition,flege1987production}; %These studies often compare the speech of bilinguals to monolingual norms, and view the influence as changes to long-term memory representations as a result of acquiring an L2 \citep{guion2003vowel,caramazza1973acquisition,flege1987production}; 
(ii) When both languages of a bilingual speaker are co-activated, by comparing speakers' mixed-language speech to their own norms while using a single language \citep[e.g.][]{grosjean1994going, bullock2009trying,elias2017effects, simonet2014phonetic}. %These compare productions during mixed-language use to participants' own productions while using a single language. The resulting influence is variously thought to involve short-term memory, online processing costs, language mode, and context-awareness. 
The present study is concerned with the latter. In production, this kind of influence has been variously termed \textit{transfer, drift, accommodation,} and \textit{interference}-- we use these interchangeably here to indicate any interaction between two sets of phonetic norms.\\

Existing research on the phonetic effects of mixed-language production has largely focused on a limited set of phonologically related languages. A majority of these studies use temporal properties of consonants (in particular, voice onset time or VOT) to measure transfer. However, the reported results vary greatly across studies, and appear to be contingent upon both language-specific features and language experience of the participants. This suggests that data from a wider variety of
%elaborate on 'great variation'-- differences between language pairs, plus sociolinguistic situation (effects of language experience etc)-- clearer. 'Types of features'--separate line?- Brief mention - discursive factors in VOT. 
populations, language pairs,  and phonetic features needs to be considered in order to make meaningful generalizations. Widespread multilingualism in the Indian subcontinent suggests that phonetic
%line--rephrase
behavior in these populations can be particularly valuable towards understanding the nature of such cross-language interactions, as they are likely to reflect real-world experience with mixed-language processing. However, there is no work yet on short-term phonetic transfer in the Indian subcontinent, or in any Indo-Aryan language. \\

The present study examines phonetic transfer between Bengali and English in a group of highly proficient bilingual speakers in India. We measure spectral properties: F1 and F2, of two English vowels to ascertain if L1 influence on L2 increases during mixed-language use, relative to a participant's unilingual baseline production of L2. Mixed-language data is elicited in two switching paradigms: cued picture-naming and code-switching. We compare these results to assess if differences between the paradigms independently influence the outcome of phonetic interaction. The results demonstrate a dynamic effect of transfer on L2 vowels during mixed-language production, but no difference in the extent of transfer between the two paradigms. Findings are discussed in light of recent proposals about asymmetries in short-term phonetic interaction and the role of connected speech in introducing discursive factors to transfer studies.\\

The rest of this section is structured as follows: \ref{bengali_english_in_india} provides background information on Bengali and Indian English focusing on vowel systems. \ref{causes} discusses two potential sources of transfer during language processing-- global co-activation of multiple languages vs the local act of switching between languages. Given the highly multilingual setting that characterizes this population, we note that the former is expected to be a constant feature and therefore any observed phonetic interaction is better understood as resulting from the latter. The design of the present study is motivated on the basis of this discussion. A logical consequence of this model is that such phonetic interaction must be highly localized. \ref{duration} elaborates on this by reviewing literature on the duration and `reversiblity' of transfer effects. A recurring pattern that emerges from existing research on short-term phonetic transfer is that both languages of the bilingual speaker are not equally affected, although the nature of these reported differences varies greatly across studies. In \ref{asymmetries}, we review some of these findings, focusing on proposed sources of asymmetries, in order to highlight the factors that mediate such phonetic interaction (or fail to do so). We argue that these results point to an urgent need for expanding the range of language pairs and sound categories examined in order to make meaningful generalizations about the nature of short-term phonetic transfer. Following from this, \ref{asymmetry between sounds} discusses reported asymmetries between different sound categories of a single language, and develops the hypotheses of the present study on the basis of these. In \ref{paradigms}, we motivate a secondary aim of this study: to verify the postulated differences between two paradigms for eliciting short-term phonetic interaction in an experimental setup, namely cued picture-naming and code-switching. \ref{questions_and_hypotheses} summarizes the research questions and hypotheses.

@@

	o-\textipa{\:t}a \quad \textipa{\ae}k-\textipa{\:t}a \quad kalo [target word]\\
	that-CL one-CL black [target word]\\
	that is a black [target word]\\
